Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-09-28 Manuscript.Rmd
Draft up to discussion ready, plus position of figures at the end

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey^1^ & Joseph T. McGuire^1^"
output:
  html_document: default
  word_document: 
    reference_docx: Style_reference.docx
  pdf_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

  Previous meta-analytic work has provided important insights on the psychological phenomena attributed to mPFC subregions. Independent coordinate-based meta-analyses have found mirroring spatial coverage in mPFC for valuation [@Bartra2013] and DMN [@Laird2009]. Whole brain analyses have shown that while more ventral mPFC regions are increasingly associated with reward-related elements of decision making, vmPFC strongly coactivates with DMN regions across the literature, with no clear topographical distinction between these phenomena [@DelaVega2016]. More focused meta-analytic work aimed at disentangling DMN from subjective value has determined that these systems are inseparable within mPFC [@Acikalin2017]. These findings highlight the multifaceted embedding of psychological constructs in subregions of mPFC, and suggest that DMN and valuation might indeed be subserved by the same system. However, recent activation-based meta-analytic work has demonstrated considerable domain specificity in this area [@Kragel2018]. While informative, a key limitation in distinguishing constructs through this approach is that meta-analyses often rely on information derived from group-averaged data. Averaging has traditionally been beneficial in identifying tendencies in brain function and organization when faced with short scanning sessions per subject. However, it has been shown that mPFC is subject to considerable idiosyncratic cortical folding [@Zilles2013] and inter-subject functional variability [@Mueller2013] compared to the rest of the brain, thus making group-averaging studies difficult to interpret.

  Seeking to remedy these shortcomings, recent work has prescribed relevance to the analysis of single subjects in fMRI. Individual analyses of fMRI data have identified idiosyncratic, reliable, and valid functional organization that would otherwise be blurred in aggregative estimations [@Laumann2015; @Gordon2017; @Gratton2018]. Furthermore, subject-specific network arrangements have been found to predict behavioral characteristics [@Kong2018]. In regards to DMN, the trend of individualized analyses has led to finer idiosyncratic subdivisions of the DMN through careful selection of seed-based functional connectivity [@Braga2017]. It is thus possible that the indisitinguishable overlap of DMN and valuation effects can be attributed to a lack of spatial resolution that is better understood at the individual level. With this in mind, an important first step in disentangling these phenomena would be to determine the degree of topological heterogeneity of the DMN present within mPFC.

@@

We performed a number of comparisons among partitions. First, we computed the degree of agreement between SP and modularity per subject to validate the resulting functional patterns. Modularity and SP show overfitting and underfitting tendencies, respectively, in their community detection performance in a diverse number of network types [@Ghasemian2018]. Their alignment would thus give us confidence in the validity of the resulting topologies. Next, we compared the functional organization derived from the whole time series among all subjects, and report the mean RI for the group. We then estimate the level of agreement between scans taken across days (i.e. pairs of daily left-right and right-left phase encoding sessions). If the functional organization estimated by SP is indeed subject-specific, we should see higher agreement for the same subject across days than either day compared to every other subject. We then extend this idea by computing the overall agreement across all sessions. Similarly, we expect that sessions within subject will show higher agreement than between subjects. Finally, we perform a sliding window analysis (20 min windows, 1 min increments, mean number of windows per subject = 37) 8) comparing each window's functional pattern against the clustering derived from the whole subject's data. This approach should give us a sense of the dynamics of the functional organization of mPFC during rest. We complement this analysis by calculating the proportion of times each surface vertex was affiliated with the DMN community across time, as well as computing the flexibility of each node (i.e. the number of times each vertex changed affiliation across time). Given that the functional heterogeneity of brain areas is unevenly distributed across the brain [@Mueller2013], we also performed these comparisons for PCC and mPFC separately. Given what we know about these general regions, we expected PCC to be well aligned across participants, and stable within subjects. On the other hand, we hypothesized low agreement of mPFC partitions across participants, but with considerable stability across days and sessions within subject. We compared the agreement within subjects against across subjects by computing the ratio of the mean RI within and between subjects. Ratios close to one denote good alignment across subjects, while ratios significantly higher than one suggest that partitions per subject were more stable than what you would expect against every other subject.

@@

A potential downfall of the present approach is that it can be computationally demanding, while more direct seed-based functional connectivity methods could provide sufficient approximations of an individual's DMN orgnization. We examined this issue by estimating mPFC functional maps from its vertex-wise functional coupling with the mean PCC activation. For each subject's concatenated day 1 and day 2 scans, we computed the mean time series of all vertices contained within the PCC region of a standard DMN arrangement [@Yeo2011], and correlated it (Pearons) with the time series of every mPFC vertex from our ROIs. The resulting day 1 and day 2 mPFC maps were then correlated within and between subjects in three ways: 1) between correlation-based maps; 2) between community-based maps; and 3) between correlation- and community-based maps. These spatial correlations are aimed to determine the day-to-day stability of each approach, as well as their level of agreement. These methods should produce somewhat similar arrangements, but the one displaying the highest within-subject agreement across days should be preferred (for example, see [@Kong2018; @Mueller2013]). We thus compared the mean within-subject inter-day correlation coefficients of each method through a paired permutation analysis (5000 iterations).

@@

```{r Meta-analysis ROIs, out.width = "60%", fig.align = "center", fig.cap= "Figure 1. Template atlas from Glasser et al. (2016) (top), as well as the common and unique brain regions associated with DMN and subjective value (bottom). Overlapping ROIs (in blue) were defined as the common brain regions that were significantly present in their respective literatures, while a single DMN-specific region (7m, in red) was identified from a permutation analysis comparing these literatures."}

include_graphics("./Images/ROIS2.png")

@@

```{r Template partition (Fig. 2), out.width="60%", fig.align = "center", fig.cap = "Figure 2. Brain partition for an example subject (100307). Fiedler vector values (top) are mapped onto the brain surface, dividing it into positive and negative communities. The bottom brain shows the binarized Fiedler vector, with yellow areas denoting the DMN community (as indicated by coverage of area 7m)."}

#include_graphics(c("./Images/newPattern.png", "./Images/newPattern_binary.png"))
include_graphics("./Images/newPattern_both.png")
```

@@

``` {r Partition comparison: all cross-method agreement (Fig. 3), fig.align="center", fig.width=3.5, fig.height=3.5, echo=FALSE, message = FALSE, cache = TRUE, fig.cap = "Figure 3. Distribution of ARI indices when comparing modularity and spectral partitioning. The distribution shows overall high agreement. (Will probably remove this figure)"}
# Cross-method
# Don't worry about plotting, since that's mainly for a general visualization
RI_all_within <- comparePartitions(Data = Summaries, Index = "RI", MOI = c("FiedlerBinary","Membership"), nSubjects = length(Summaries), subjNames = subjList)

RI_within_plot <- ggplot(data = RI_all_within, aes(Index)) + 
        geom_density(fill = Cols[3]) + 
        xlim(0,1) + 
        labs(x = "Adj. Rand Index") +
        theme_classic()
RI_within_plot
```

``` {r Partition comparison: all 2, echo=FALSE, message = FALSE, cache = TRUE}
# Across subjects
# Note: the upper and lower triangles can totally be combined before plotting, instead of adding one corrplot over another. Might work on that eventually.
RI_all_between <- comparePartitions(Data = Summaries, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
RI_all_between_lowtri <- RI_all_between[lower.tri(RI_all_between)]

@@

``` {r Partition comparison: all 2 (Fig. 4), fig.align="center", out.width = "40%", fig.cap = "Figure 4. Similarity matrix showing ARI values among all subjects for PCC (lower triangle) and mPFC (upper triangle) separately. PCC is well aligned across individuals, while mPFC shows clear topographical heterogeneity.", echo=FALSE, message = FALSE, cache = TRUE}

# par(mfrow = c(1,2))
# corrplot(RI_all_between,
#          is.corr = F, 
#          method = "color", 
#          tl.pos = "n", 
#          tl.col = "black", 
#          cl.lim = c(0,1),
#          cl.pos = "b",
#          cl.length = 3)
# mtext("Subjects", side = 2, line = -4.5, cex = 1.5)
# mtext("Subjects", side = 3, line = 4.5, cex = 1.5)
par(xpd=T)
corrplot(RI_PCC,
         is.corr = F, 
         method = "color",
         type = "lower", 
         tl.pos = "n", 
         tl.cex = 0.1, 
         cl.lim = c(0,1),
         cl.length = 3,
         addgrid.col = "grey99",
         cl.pos = "r",
         mar = c(1,1,2,0))
corrplot(RI_PFC, 
         is.corr = F, 
         method = "color", 
         type = "upper", 
         tl.pos = "n", 
         tl.cex = 0.1,
         cl.pos = "n",
         addgrid.col = "grey90",
         mar = c(1,1,2,0), 
         add = T)
segments(0.5,0.5,0.5,100.5, lwd = 2)
segments(0.5,0.5,100.5,0.5, lwd = 2)
segments(0.5,100.5,100.5,100.5, lwd = 2)
segments(100.5,0.5,100.5,100.5, lwd = 2)
mtext(expression(bold("Subject Partitions")), line = 3)
mtext(expression(bold("Subject Partitions")), side = 2, line = -0.8)
mtext(expression(bold("mPFC")), line = 2, at = par("usr")[1]+0.73*diff(par("usr")[1:2]), col = "steelblue4")
mtext(expression(bold("PCC")), side = 1, line = 4, at = par("usr")[1]+0.18*diff(par("usr")[1:2]), col = "steelblue4")

@@

``` {r Partition comparison: halves ratios (Fig. 6), fig.align="center", fig.width=3, fig.height=3, fig.cap = "Figure 6. mPFC displays higher within-subject agreement than between subjeccts, as evidenced by a high within/between subject mean ARI. PCC also shows some level of topographical idiosyncrasy (ratio > 1), but significantly lower than mPFC.", echo=FALSE, message = FALSE, cache = TRUE}
## Now compare all possible pairs of subjects
# Compare all halves with RI
RI_all_between_halves <- comparePartitions(Data = Summaries_halves, Index = "RI", nSubjects = length(Summaries_halves), subjNames = subjList_halves)

@@

# plot
ggplot(data = Halves_RIRatios_summary, aes(Region, RI, fill = Region)) + 
  geom_boxplot(show.legend = F) +
  scale_fill_manual(values = Cols[seq(2)]) + 
  ylim(0, 10) +
  geom_hline(yintercept = 1) +
  labs(y = "ARI Ratio", x = "") + 
  theme_classic(base_size = 14)

@@

``` {r Partition comparison: session ratios (Fig. 8), fig.align="center", fig.width=3, fig.height=3, fig.cap = "Figure 8. Whole group ratio between within- and between-subject ARI for mPFC and PCC based on sessions. Consistent with the daily analysis, both regions show idiosyncrasy, with mPFC being significantly more subject-specific.", echo=FALSE, message = FALSE, cache = TRUE}
## Brain storming:
## Within-subject comparisons would be more interesting? Like, comparing the mean RI within-between per subject
## Perhaps a more interresting idea would be to generate an index of variance within/between subject
## These are just attempts at that are not perfect:

@@

# Stats
Sess_mPFCvPCC_permTest <- permute(Sess_RIRatios_PFC, Sess_RIRatios_PCC, paired = T)
Sess_mPFCvPCC_ES <- cohenD(Sess_RIRatios_PFC, Sess_RIRatios_PCC)

ggplot(data = Sess_RIRatios_summary, aes(Region, RI, fill = Region)) + 
  geom_boxplot(show.legend = F) +
  scale_fill_manual(values = Cols[seq(2)]) + 
  ylim(0, NA) +
  geom_hline(yintercept = 1) +
  labs(y = "ARI Ratio", x = "") + 
  theme_classic(base_size = 14)
```

@@

``` {r Stability per subject: Overall (Fig. 9), fig.align="center", fig.width=5, fig.height=3, fig.cap = "Figure 9. ARI between communities computed at each sliding epoch and the original partition for all subjects (left), as well as the distribution of ARI standard deviations per subject (right).", echo=FALSE, message = FALSE, cache = TRUE}
# Load files
setwd('./Sliding_window/')
temp <- list.files(pattern = '*Comparisons.csv')
SC <- lapply(temp, read.csv2)
temp <- list.files(pattern = '*Values.csv')
SVals <- lapply(temp, read.csv2)

@@

grid.arrange(SC_plot, SC_sd_plot, ncol = 2)

@@

``` {r Noise Effects: tSNR Comparison (Fig. 12), fig.align="center", fig.width=3, fig.height=3, fig.cap = "Figure 12. Mean tSNR for mPFC and PCC across participants. This measure was derived from the minimally preprocessed fMRI data from the HCP, and thus shows the base level of reltive noise in each region.", echo=FALSE, message = FALSE, cache = TRUE}
## Putative steps to analyze the SNR of the full time series (PCC vs mPFC)
# Flexibility vs SNR
# Are vertices with lower SNR switching communities more often?
SNR_vs_Flex <- sapply(Summaries, function(data) cor(data$varCoeff, data$slideFlexibility))

@@

# Get the mean SNR for each region and compare overall levels 
SNR_meanPCC <- sapply(tempComm_PCC, function(data) mean(data$varCoeff))
SNR_meanPFC <- sapply(tempComm_PFC, function(data) mean(data$varCoeff))
SNR_meanComp <- list(Values = data.frame(Region = c(rep("PCC", nSubj), rep("mPFC", nSubj)), SNR = c(SNR_meanPCC, SNR_meanPFC)),
                     Comparison = permute(SNR_meanPCC, SNR_meanPFC)$Pval,
                     ES = cohenD(SNR_meanPCC, SNR_meanPFC))
rm(SNR_meanPCC, SNR_meanPFC)

# Plot the mPFC SNR vs PCC SNR for each subject (one figure containing all subjects)
snrmeanComp_plot <- ggplot(data = SNR_meanComp$Values, aes(Region, SNR, fill = Region)) +
  geom_boxplot(show.legend = F) +
  scale_fill_manual(values = Cols[seq(2)]) +
  labs(x = "", y = "tSNR") +
  theme(text = element_text(size=35)) +
  theme_classic()
snrmeanComp_plot
```

@@

``` {r Noise Effects: Per-subject Fiedler by R-sqd (Fig. 13), fig.align="center", fig.width=3, fig.height=3, fig.cap = "Figure 13. Fiedler vector values are linearly related to the total variance explained by each vertex with respect to the mean DMN time series.", echo=FALSE, message = FALSE, cache = TRUE}
setwd('./Variance/')
temp <- list.files()
Variance <- lapply(temp, read.csv)

@@

``` {r Noise Effects: Mean R-sqd mPFC and PCC, fig.align="center", fig.width=6, fig.height=4, fig.cap = "Figure 14. Differences in mean R-squared values for each region per participant. The figure shows that PCC vertices tend to be more coupled with the overall dynamics of DMN across subjects.", echo=FALSE, message = FALSE}
v4
```

To further interrogate the potential effect of noise in our results, we computed the variance explained by activity from each DMN vertex on the mean time series of the DMN community. If the signal quality issues persisted after preprocessing the data, we expected that noisy vertices would have the least explicative power on the activity of the DMN community as a whole. This also implied lower $R^2$ values for mPFC. Indeed, we found that for each subject, $R^2$ values tended to be higher for PCC (`r prcntSig` percent of Wilcoxon ranksum tests across subjects had *p* values < 0.001; mean Cohen's D = `r round(mean(Rsq_PCCvsPFC_ES), digits = 2)`) (Figure 14). These results suggest that, overall, vertices in mPFC tended to be less in synchrony with the activity of the DMN community as a whole. Furthermore, the degree tSNR showed a moderate correlation with $R^2$ values across participants (mean *r* = `r round(mean(noiseCors$R2vtSNR), digits = 2)`; SD = `r round(sd(noiseCors$R2vtSNR), digits = 2)`), meaning that noisier vertices were less in synchrony with the general activity of the DMN.

Given this trend, we asked whether there is a "gradient of noise" along the z-axis of mPFC, such that more ventral vertices would both be noisier and explain the least amount of DMN variance. Here we found that while $R^2$ are somewhat correlated with dorsal-ventral placement across subjects (mean *r* = `r round(mean(noiseCors$ZvR2), digits = 2)`; SD = `r round(sd(noiseCors$ZvR2), digits = 2)`), there is no clear relationship with the z-axis across subjects (Figure 15). Specifically, low $R^2$ are evenly distributed along this axis, denoting a degree of orthogonality between these features. Conversely, tSNR showed a strong linear relationship with the position of the vertex along this axis (mean *r* = `r round(mean(noiseCors$ZvtSNR), digits = 2)`; SD = `r round(sd(noiseCors$ZvtSNR), digits = 2)`). These results contextualize the degree of synchrony of mPFC vertices in terms of potential noise. While still an important factor, it is unlikely that the degree of within-subject pattern stability observed here could be due mostly to highly structured mPFC noise.

``` {r Noise Effects: R-sqd vs tSNR (Fig. 15), fig.align="center", fig.width=5, fig.height=3, fig.cap = "Figure 15. Vertex tSNR values across all subjects are correlated with their position along the z-axis in mPFC, with vmPFC showing the lowest tSNR (left). However, R-square values do not show this strong relationship (right), suggesting that the coupling of each vertex is not necessarily influenced by signal noise.", echo=FALSE, message = FALSE, cache = TRUE}
# grid.arrange(v6, v7, ncol = 2)
# par(mfrow = c(1,2))
# smoothScatter(allVariance[allVariance$Region == "mPFC",]$Rsquared, allVariance[allVariance$Region == "mPFC",]$tempZ,
#               xlab = "R-squared Value",
#               ylab = "Z-Coordinate")
# 
# smoothScatter(allVariance[allVariance$Region == "mPFC",]$varCoeff, allVariance[allVariance$Region == "mPFC",]$tempZ,
#               xlab = "Coefficient of Variation (SNR)",
#               ylab = "")

@@

The computational demands of our approach escalate quickly as the number of vertices increases. We thus explored the possible advantage to using graph-theoretic approaches on all possible network connections over a simple seed-based functional connectivity analysis from PCC. Given that the spatial distribution of the DMN is considerably stable within individuals [@Laumann2015; @Gordon2017; @Braga2017; @Kong2018], we examined whether maps generated with spectral partitioning were more similar per subject across days than those computed from seed-based correlations. First, we computed the mean time series of all the vertices contained in the DMN PCC ROI from Yeo et al. (2011) parcellation, and correlated it with the activity of every mPFC vertex from our ROIs. We then performed pairwise spatial Pearson correlations among maps computed for each day and method from all subjects (Figure 16).

Figure 17 shows the overall degree of similarity across days for patterns dervied using community detection, seed-based correlation, and across methods (i.e. day one community detection versus day two seed-based correlation). While both approaches are very consistent, community detection displayed a significantly higher ARI across days than a simple correlation (paired permutation *p* < 0.001; Cohen's D = `r round(prepostComp_ES, digits = 2)`). It is worth noting that the cross-method agreement is fair, meaning that both approaches show similar topological features. These findings suggest that graph-theoretic community detection algorithms provide a better way to detect stable functional topologies, an important factor when faced with the levels of noise present in mPFC.

``` {r Correlation vs community detection (Fig. 16), fig.align="center", fig.width = 4, fig.height = 4, fig.cap = "Figure 16. Correlation matrix comparing the inter-day spatial stability of maps derived from seed-based functional connectivity (PCC-based) and the Fiedler vector (10 example subjects). The top-left quadrant represents seed-based FC maps, and the bottom-right the Fiedler vector.", echo=FALSE, message = FALSE, cache = TRUE}
corrplot(allHalfCombs, 
         #title = "Comparing All Community and Correlation mPFC Partitions \n Across All Subjects",
         method = "color", 
         #tl.col = "black",
         #tl.cex = 0.6,
         addgrid.col = "grey90",
         tl.pos = "n",
         cl.pos = "r",
         cl.length = 3,
         mar = c(1,2,2,0))
segments(0.5,0.5,0.5,40.5, lwd = 2)
segments(0.5,0.5,40.5,0.5, lwd = 2)
segments(0.5,40.5,40.5,40.5, lwd = 2)
segments(40.5,0.5,40.5,40.5, lwd = 2)
segments(0.5,20.5,40.5,20.5, lwd = 2)
segments(20.5,0.5,20.5,40.5, lwd = 2)
mtext(expression(bold("Subject Partitions (Day 1 and Day 2)")), line = 3, at = par("usr")[1]+0.45*diff(par("usr")[1:2]))
mtext(expression(bold("Subject Partitions (Day 1 and Day 2)")), side = 2, line = 3)
mtext(expression(bold("Fiedler")), line = 1.5, at = par("usr")[1]+0.68*diff(par("usr")[1:2]))
mtext(expression(bold("Correlation")), line = 1.5, at = par("usr")[1]+0.23*diff(par("usr")[1:2]))
mtext(expression(bold("Fiedler")), side = 2, line = 2, at = par("usr")[1]+0.25*diff(par("usr")[1:2]))
mtext(expression(bold("Correlation")), side = 2, line = 2, at = par("usr")[1]+0.68*diff(par("usr")[1:2]))

``` {r Correlation vs community detection: comparison plot (Fig. 17), fig.align="center", fig.width=3, fig.height=3, fig.cap = "Figure 17. Day 1 vs Day 2 oefficients of spatial correlations for every method, as well as between methods. Community detection through spectral partitioning provides the most stable inter-day estimates, even though both methods show good level of agreement.", echo=FALSE, message = FALSE}
# Day 1 vs Day 2 Spatial Correlation Comparison
prepostComp_plot <- ggplot(data = prepostComp, aes(Approach, Correlation, fill = Approach)) +
  geom_boxplot(show.legend = F) +
  labs(x = "", y = "Correlation Coefficient") +
  ylim(0,1) +
  theme(text = element_text(size=35)) +
  theme_classic()
prepostComp_plot
```

@@

Finally, we examined whether these topological features were defined by the structure of the brain by correlating each subject's Fiedler vector with their structural MRI data. We found that the Fiedler vector is uncorrelated with myelin density (mean correlation = `r round(meanMyelin, digits = 2)`; SD = `r round(sdMyelin, digits = 2)`), cortical folding (mean correlation = `r round(meanCurv, digits = 2)`; SD = `r round(sdCurv, digits = 2)`), and surface thickness (mean correlation = `r round(meanThick, digits = 2)`; SD = `r round(sdThick, digits = 2)`). While these results suggest that each subject's functional topology is mostly independent of brain structure, we will perform more detailed analyses of this in the future (i.e. are there any outstanding structural features in a subject that match their unique topology?).

@@

## Acknowledgements

Data were provided [in part] by the Human Connectome Project, WU-Minn Consortium (Principal Investigators: David Van Essen and Kamil Ugurbil; 1U54MH091657) funded by the 16 NIH Institutes and Centers that support the NIH Blueprint for Neuroscience Research; and by the McDonnell Center for Systems Neuroscience at Washington University.

>>>>

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey^1^ & Joseph T. McGuire^1^"
output:
  pdf_document: default
  html_document: default
  word_document: 
    reference_docx: Style_reference.docx
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

  Previous meta-analytic work has provided important insights on the psychological phenomena attributed to mPFC subregions. Independent coordinate-based meta-analyses have found mirroring spatial coverage in mPFC for valuation [@Bartra2013] and DMN [@Laird2009]. Whole brain analyses have shown that while more ventral mPFC regions are increasingly associated with reward-related elements of decision making, vmPFC strongly coactivates with DMN regions across the literature, with no clear topographical distinction between these phenomena [@DelaVega2016]. More focused meta-analytic work aimed at disentangling DMN from subjective value has determined that these systems are inseparable within mPFC [@Acikalin2017]. These findings highlight the multifaceted embedding of psychological constructs in subregions of mPFC, and suggest that DMN and valuation might indeed be subserved by the same system. However, recent activation-based meta-analytic work has demonstrated considerable domain specificity in this area [@Kragel2018]. While informative, a key limitation in distinguishing constructs through this approach is that meta-analyses often rely on information derived from group-averaged data. Averaging has traditionally been beneficial in identifying tendencies in brain function and organization when faced with short scanning sessions per subject. However, it has been shown that mPFC is subject to considerable idiosyncratic cortical folding [@Zilles2013] and inter-subject functional variability [@Mueller2013a] compared to the rest of the brain, thus making group-averaging studies difficult to interpret.

  Seeking to remedy these shortcomings, recent work has prescribed relevance to the analysis of single subjects in fMRI. Individual analyses of fMRI data have identified idiosyncratic, reliable, and valid functional organization that would otherwise be blurred in aggregative estimations [@Laumann2015; @Gordon2017; @Gratton2018a]. Furthermore, subject-specific network arrangements have been found to predict behavioral characteristics [@Kong2018]. In regards to DMN, the trend of individualized analyses has led to finer idiosyncratic subdivisions of the DMN through careful selection of seed-based functional connectivity [@Braga2017]. It is thus possible that the indisitinguishable overlap of DMN and valuation effects can be attributed to a lack of spatial resolution that is better understood at the individual level. With this in mind, an important first step in disentangling these phenomena would be to determine the degree of topological heterogeneity of the DMN present within mPFC.

@@

We performed a number of comparisons among partitions. First, we computed the degree of agreement between SP and modularity per subject to validate the resulting functional patterns. Modularity and SP show overfitting and underfitting tendencies, respectively, in their community detection performance in a diverse number of network types [@Ghasemian2018]. Their alignment would thus give us confidence in the validity of the resulting topologies. Next, we compared the functional organization derived from the whole time series among all subjects, and report the mean RI for the group. We then estimate the level of agreement between scans taken across days (i.e. pairs of daily left-right and right-left phase encoding sessions). If the functional organization estimated by SP is indeed subject-specific, we should see higher agreement for the same subject across days than either day compared to every other subject. We then extend this idea by computing the overall agreement across all sessions. Similarly, we expect that sessions within subject will show higher agreement than between subjects. Finally, we perform a sliding window analysis (20 min windows, 1 min increments, mean number of windows per subject = 37) 8) comparing each window's functional pattern against the clustering derived from the whole subject's data. This approach should give us a sense of the dynamics of the functional organization of mPFC during rest. We complement this analysis by calculating the proportion of times each surface vertex was affiliated with the DMN community across time, as well as computing the flexibility of each node (i.e. the number of times each vertex changed affiliation across time). Given that the functional heterogeneity of brain areas is unevenly distributed across the brain [@Mueller2013a], we also performed these comparisons for PCC and mPFC separately. Given what we know about these general regions, we expected PCC to be well aligned across participants, and stable within subjects. On the other hand, we hypothesized low agreement of mPFC partitions across participants, but with considerable stability across days and sessions within subject. We compared the agreement within subjects against across subjects by computing the ratio of the mean RI within and between subjects. Ratios close to one denote good alignment across subjects, while ratios significantly higher than one suggest that partitions per subject were more stable than what you would expect against every other subject.

@@

A potential downfall of the present approach is that it can be computationally demanding, while more direct seed-based functional connectivity methods could provide sufficient approximations of an individual's DMN orgnization. We examined this issue by estimating mPFC functional maps from its vertex-wise functional coupling with the mean PCC activation. For each subject's concatenated day 1 and day 2 scans, we computed the mean time series of all vertices contained within the PCC region of a standard DMN arrangement [@Yeo2011], and correlated it (Pearons) with the time series of every mPFC vertex from our ROIs. The resulting day 1 and day 2 mPFC maps were then correlated within and between subjects in three ways: 1) between correlation-based maps; 2) between community-based maps; and 3) between correlation- and community-based maps. These spatial correlations are aimed to determine the day-to-day stability of each approach, as well as their level of agreement. These methods should produce somewhat similar arrangements, but the one displaying the highest within-subject agreement across days should be preferred (for example, see [@Kong2018; @Mueller2013a]). We thus compared the mean within-subject inter-day correlation coefficients of each method through a paired permutation analysis (5000 iterations).

@@

@@

@@

``` {r Partition comparison: all 2, echo=FALSE, message = FALSE, cache = TRUE}
# Cross-method
# Don't worry about plotting, since that's mainly for a general visualization
RI_all_within <- comparePartitions(Data = Summaries, Index = "RI", MOI = c("FiedlerBinary","Membership"), nSubjects = length(Summaries), subjNames = subjList)

# Across subjects
# Note: the upper and lower triangles can totally be combined before plotting, instead of adding one corrplot over another. Might work on that eventually.
RI_all_between <- comparePartitions(Data = Summaries, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
RI_all_between_lowtri <- RI_all_between[lower.tri(RI_all_between)]

@@

@@

``` {r Partition comparison: halves ratios, echo=FALSE, message = FALSE, cache = TRUE}
## Now compare all possible pairs of subjects
# Compare all halves with RI
RI_all_between_halves <- comparePartitions(Data = Summaries_halves, Index = "RI", nSubjects = length(Summaries_halves), subjNames = subjList_halves)

@@

@@

``` {r Partition comparison: session ratios, echo=FALSE, message = FALSE, cache = TRUE}
## Brain storming:
## Within-subject comparisons would be more interesting? Like, comparing the mean RI within-between per subject
## Perhaps a more interresting idea would be to generate an index of variance within/between subject
## These are just attempts at that are not perfect:

@@

# Stats
Sess_mPFCvPCC_permTest <- permute(Sess_RIRatios_PFC, Sess_RIRatios_PCC, paired = T)
Sess_mPFCvPCC_ES <- cohenD(Sess_RIRatios_PFC, Sess_RIRatios_PCC)
```

@@

``` {r Stability per subject: Overall, echo=FALSE, message = FALSE, cache = TRUE}
# Load files
setwd('./Sliding_window/')
temp <- list.files(pattern = '*Comparisons.csv')
SC <- lapply(temp, read.csv2)
temp <- list.files(pattern = '*Values.csv')
SVals <- lapply(temp, read.csv2)

@@

@@

``` {r Noise Effects: tSNR Comparison, echo=FALSE, message = FALSE, cache = TRUE}
## Putative steps to analyze the SNR of the full time series (PCC vs mPFC)
# Flexibility vs SNR
# Are vertices with lower SNR switching communities more often?
SNR_vs_Flex <- sapply(Summaries, function(data) cor(data$varCoeff, data$slideFlexibility))

@@

# Get the mean SNR for each region and compare overall levels 
SNR_meanPCC <- sapply(tempComm_PCC, function(data) mean(data$varCoeff))
SNR_meanPFC <- sapply(tempComm_PFC, function(data) mean(data$varCoeff))
SNR_meanComp <- list(Values = data.frame(Region = c(rep("PCC", nSubj), rep("mPFC", nSubj)), SNR = c(SNR_meanPCC, SNR_meanPFC)),
                     Comparison = permute(SNR_meanPCC, SNR_meanPFC)$Pval,
                     ES = cohenD(SNR_meanPCC, SNR_meanPFC))
rm(SNR_meanPCC, SNR_meanPFC)
```

@@

``` {r Noise Effects: Per-subject Fiedler by R-sqd, fig.align="center", fig.width=3, fig.height=3, fig.cap = "FORMER Figure 13. Fiedler vector values are linearly related to the total variance explained by each vertex with respect to the mean DMN time series.", echo=FALSE, message = FALSE, cache = TRUE}
setwd('./Variance/')
temp <- list.files()
Variance <- lapply(temp, read.csv)

@@

To further interrogate the potential effect of noise in our results, we computed the variance explained by activity from each DMN vertex on the mean time series of the DMN community. If the signal quality issues persisted after preprocessing the data, we expected that noisy vertices would have the least explicative power on the activity of the DMN community as a whole. This also implied lower $R^2$ values for mPFC. Indeed, we found that for each subject, $R^2$ values tended to be higher for PCC (`r prcntSig` percent of Wilcoxon ranksum tests across subjects had *p* values < 0.001; mean Cohen's D = `r round(mean(Rsq_PCCvsPFC_ES), digits = 2)`) (Figure 13). These results suggest that, overall, vertices in mPFC tended to be less in synchrony with the activity of the DMN community as a whole. Furthermore, the degree tSNR showed a moderate correlation with $R^2$ values across participants (mean *r* = `r round(mean(noiseCors$R2vtSNR), digits = 2)`; SD = `r round(sd(noiseCors$R2vtSNR), digits = 2)`), meaning that noisier vertices were less in synchrony with the general activity of the DMN.

Given this trend, we asked whether there is a "gradient of noise" along the z-axis of mPFC, such that more ventral vertices would both be noisier and explain the least amount of DMN variance. Here we found that while $R^2$ are somewhat correlated with dorsal-ventral placement across subjects (mean *r* = `r round(mean(noiseCors$ZvR2), digits = 2)`; SD = `r round(sd(noiseCors$ZvR2), digits = 2)`), there is no clear relationship with the z-axis across subjects (Figure 14). Specifically, low $R^2$ are evenly distributed along this axis, denoting a degree of orthogonality between these features. Conversely, tSNR showed a strong linear relationship with the position of the vertex along this axis (mean *r* = `r round(mean(noiseCors$ZvtSNR), digits = 2)`; SD = `r round(sd(noiseCors$ZvtSNR), digits = 2)`). These results contextualize the degree of synchrony of mPFC vertices in terms of potential noise. While still an important factor, it is unlikely that the degree of within-subject pattern stability observed here could be due mostly to highly structured mPFC noise.

@@

The computational demands of our approach escalate quickly as the number of vertices increases. We thus explored the possible advantage to using graph-theoretic approaches on all possible network connections over a simple seed-based functional connectivity analysis from PCC. Given that the spatial distribution of the DMN is considerably stable within individuals [@Laumann2015; @Gordon2017; @Braga2017; @Kong2018], we examined whether maps generated with spectral partitioning were more similar per subject across days than those computed from seed-based correlations. First, we computed the mean time series of all the vertices contained in the DMN PCC ROI from Yeo et al. (2011) parcellation, and correlated it with the activity of every mPFC vertex from our ROIs. We then performed pairwise spatial Pearson correlations among maps computed for each day and method from all subjects (Figure 15).

Figure 16 shows the overall degree of similarity across days for patterns dervied using community detection, seed-based correlation, and across methods (i.e. day one community detection versus day two seed-based correlation). While both approaches are very consistent, community detection displayed a significantly higher ARI across days than a simple correlation (paired permutation *p* < 0.001; Cohen's D = `r round(prepostComp_ES, digits = 2)`). It is worth noting that the cross-method agreement is fair, meaning that both approaches show similar topological features. These findings suggest that graph-theoretic community detection algorithms provide a better way to detect stable functional topologies, an important factor when faced with the levels of noise present in mPFC.

@@

Finally, we coarsely examined whether these idiosyncratic topographical arrangements were defined by anatomical brain features. To do so, we correlated each subject's Fiedler vector with their respective structural MRI data. On average, we found that the Fiedler vector is uncorrelated with myelin density (mean *r* = `r round(meanMyelin, digits = 2)`; SD = `r round(sdMyelin, digits = 2)`), cortical folding (mean *r* = `r round(meanCurv, digits = 2)`; SD = `r round(sdCurv, digits = 2)`), and surface thickness (mean *r* = `r round(meanThick, digits = 2)`; SD = `r round(sdThick, digits = 2)`). While these findings point at an orthogonality between structure and functional organization, they resonate with previous work that found no strong matches between intersubject functional variability and cortical thickness, and low variability in sulcal depth within DMN [@Mueller2013a].

@@

\newpage
## Acknowledgements

We would like to thank Dan Sussman for initial guidance on community detection and evaluation methods. Data were provided [in part] by the Human Connectome Project, WU-Minn Consortium (Principal Investigators: David Van Essen and Kamil Ugurbil; 1U54MH091657) funded by the 16 NIH Institutes and Centers that support the NIH Blueprint for Neuroscience Research; and by the McDonnell Center for Systems Neuroscience at Washington University.
